---
title: "Homework 6"
author: Yujie Li
date: 2024-12-01
output: 
  html_document:
    toc: true 
    toc_float: true 
   
output: github_document 
---


## Problem 1

```{r, echo = FALSE, message = FALSE}
library(dplyr)
library(broom)
library(ggplot2)
library(rnoaa)
```

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```


```{r}
set.seed(123)

bootstrap_function <- function(data) {
  sample_data <- data %>% sample_frac(size = 1, replace = TRUE)
  model <- lm(tmax ~ tmin, data = sample_data)
  
  r_squared <- broom::glance(model)$r.squared
  coefs <- broom::tidy(model)
  log_beta_product <- log(coefs$estimate[1] * coefs$estimate[2])
  
  c(r_squared = r_squared, log_beta_product = log_beta_product)
}

bootstrap_results <- replicate(
  5000,
  bootstrap_function(weather_df),
  simplify = "matrix"
)
bootstrap_df <- as.data.frame(t(bootstrap_results))
colnames(bootstrap_df) <- c("r_squared", "log_beta_product")
```

```{r}
ci_r_squared <- quantile(bootstrap_df$r_squared, probs = c(0.025, 0.975))
ci_log_beta <- quantile(bootstrap_df$log_beta_product, probs = c(0.025, 0.975))

ggplot(bootstrap_df, aes(x = r_squared)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black", alpha = 0.7) +
  geom_vline(xintercept = ci_r_squared, linetype = "dashed", color = "red") +
  labs(title = "Bootstrap Distribution of R-Squared",
       x = "R-Squared", y = "Frequency") +
  theme_minimal()

ggplot(bootstrap_df, aes(x = log_beta_product)) +
  geom_histogram(bins = 30, fill = "orange", color = "black", alpha = 0.7) +
  geom_vline(xintercept = ci_log_beta, linetype = "dashed", color = "red") +
  labs(title = "Bootstrap Distribution of Log(Beta0 * Beta1)",
       x = "Log(Beta0 * Beta1)", y = "Frequency") +
  theme_minimal()
```


## Problem 2 

```{r, echo = FALSE, message = FALSE}
library(dplyr)
library(broom)
library(purrr)
library(ggplot2)
library(tidyr)
```

```{r}
homicide_data <- read.csv("homicide-data.csv", stringsAsFactors = FALSE)

homicide_data <- homicide_data %>%
  mutate(city_state = paste(city, state, sep = ", "))

homicide_data <- homicide_data %>%
  mutate(resolved = ifelse(grepl("Closed", disposition), 1, 0))

excluded_cities <- c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")
homicide_data <- homicide_data %>%
  filter(!city_state %in% excluded_cities)

homicide_data <- homicide_data %>%
  filter(victim_race %in% c("White", "Black"))

homicide_data <- homicide_data %>%
  mutate(victim_age = as.numeric(victim_age)) %>%
  filter(!is.na(victim_age))


```
```{r}
# Identify rows with problematic victim_age
invalid_ages <- homicide_data %>%
  filter(is.na(as.numeric(victim_age)) & !is.na(victim_age))

# Display rows with invalid victim_age
print(invalid_ages)

# Remove rows with invalid victim_age or fix them manually if needed
homicide_data <- homicide_data %>%
  mutate(victim_age = as.numeric(victim_age)) %>%
  filter(!is.na(victim_age))

```



```{r}
baltimore_data <- homicide_data %>%
  filter(city_state == "Baltimore, MD")

model_baltimore <- glm(
  resolved ~ victim_age + victim_sex + victim_race,
  family = binomial(link = "logit"),
  data = baltimore_data
)

baltimore_summary <- tidy(model_baltimore, conf.int = TRUE) %>%
  mutate(
    odds_ratio = exp(estimate),
    conf.low = exp(conf.low),
    conf.high = exp(conf.high)
  )

print(baltimore_summary)
```

```{r}
city_models <- homicide_data %>%
  group_by(city_state) %>%
  nest() %>%
  mutate(
    model = map(data, ~ glm(resolved ~ victim_age + victim_sex + victim_race,
                            family = binomial(link = "logit"), data = .)),
    tidied = map(model, ~ tidy(., conf.int = TRUE) %>%
                   mutate(
                     odds_ratio = exp(estimate),
                     conf.low = exp(conf.low),
                     conf.high = exp(conf.high)
                   ) %>%
                   filter(term == "victim_sexMale"))
  ) %>%
  unnest(tidied)

city_plot <- city_models %>%
  ggplot(aes(x = reorder(city_state, odds_ratio), y = odds_ratio)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  coord_flip() +
  labs(
    title = "Adjusted Odds Ratios for Solving Homicides",
    x = "City",
    y = "Odds Ratio (Male vs Female Victims)"
  ) +
  theme_minimal()

print(city_plot)
```
## Comment on the graph
Many cities have ORs close to 1, indicating minimal gender-based differences in homicide resolution rates.
Some cities, such as Fresno, CA, and Minneapolis, MN, show ORs significantly greater than 1, implying higher resolution rates for male victims.
Conversely, cities like Long Beach, CA, and New York, NY, show ORs less than 1, suggesting female victims have higher resolution rates in these areas.


## Problem 3

```{r, echo = FALSE, message = FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(modelr)
library(purrr)
```

```{r}
# Load the dataset
birth_data <- read.csv("birthweight.csv", stringsAsFactors = FALSE)

# Convert numeric variables to factors where appropriate
birth_data <- birth_data %>%
  mutate(
    babysex = factor(babysex, levels = c(1, 2), labels = c("Male", "Female")),
    malform = factor(malform, levels = c(0, 1), labels = c("Absent", "Present")),
    frace = factor(frace, levels = c(1, 2, 3, 4, 8, 9)),
    mrace = factor(mrace, levels = c(1, 2, 3, 4, 8))
  )

# Check for missing data
missing_summary <- colSums(is.na(birth_data))
print(missing_summary)

# Impute missing data (if needed)
birth_data <- birth_data %>%
  drop_na() # Alternatively, you can impute missing values with other methods

```
```{r}
# Fit the hypothesized model
model_hypothesis <- lm(bwt ~ gaweeks + wtgain + mheight + bhead + malform + smoken, data = birth_data)

# Summarize the model
summary(model_hypothesis)

```

```{r}
# Add predictions and residuals to the dataset
birth_data <- birth_data %>%
  add_predictions(model_hypothesis, var = "fitted_values") %>%
  add_residuals(model_hypothesis, var = "residuals")

# Plot residuals vs fitted values
ggplot(birth_data, aes(x = fitted_values, y = residuals)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Residuals vs Fitted Values",
    x = "Fitted Values",
    y = "Residuals"
  ) +
  theme_minimal()

```

## Comparison to two other models
```{r}
model_length_gest <- lm(bwt ~ blength + gaweeks, data = birth_data)
model_interactions <- lm(bwt ~ bhead * blength * babysex, data = birth_data)

```


```{r}
set.seed(123)
n_splits <- 100
cv_results <- replicate(n_splits, {
  split <- split_data(birth_data, prop = 0.8)
  train <- split$train
  test <- split$test
  
  model_hypothesis <- lm(bwt ~ gaweeks + wtgain + mheight + bhead + malform + smoken, data = train)
  model_length_gest <- lm(bwt ~ blength + gaweeks, data = train)
  model_interactions <- lm(bwt ~ bhead * blength * babysex, data = train)
  
  hypothesis_rmse <- rmse(test$bwt, predict(model_hypothesis, newdata = test))
  length_gest_rmse <- rmse(test$bwt, predict(model_length_gest, newdata = test))
  interactions_rmse <- rmse(test$bwt, predict(model_interactions, newdata = test))
  
  c(hypothesis_rmse, length_gest_rmse, interactions_rmse)
})

cv_results <- t(cv_results)

cv_summary <- data.frame(
  Model = c("Hypothesized Model", "Length + Gestational Age", "Interactions"),
  Mean_RMSE = colMeans(cv_results)
)

print(cv_summary)


```

